---
title: "Assignment 7"
author: "Bishoy Sokkar"
date: "2024-03-15"
output: html_document
---

##Intro

In the following codes we will create a table of books in HTML, XML, and JSON, with a goal of mind to pull and create dataframes in R.
Loading out packages are in the next code:


```{r}
library(tidyverse)
library(rvest)
library(XML)
library(jsonlite)
```

## HTML

First I created the HTML table.
```{r}
table_HTML <- read_html("https://raw.githubusercontent.com/sokkarbishoy/DATA607/main/Assignment%207%20HTML%20table")
```

Second we load into a dataframe in R using the following code. 

```{r}
Books_table_HTML <- table_HTML %>% 
  html_node("table") %>%
  html_table(header = TRUE, fill = TRUE)
Books_table_HTML
```

##XML 
same as we did in the first part, we store out XML code in Github and read it on R
```{r}
library(xml2)
library(xmlconvert)
url <- "https://raw.githubusercontent.com/sokkarbishoy/DATA607/main/Books%20.XML"

xmltable <- (read_xml(url))
xmltable

xml_data <- xmlParse(xmltable)
```

```{r}
titles <- xpathSApply(xml_data, "//book/title", xmlValue)
authors <- xpathSApply(xml_data, "//book/authors", xmlValue)
years <- xpathSApply(xml_data, "//book/publication_year", xmlValue)
summaries <- xpathSApply(xml_data, "//book/summary", xmlValue)
main_themes <- xpathSApply(xml_data, "//book/main_themes", xmlValue)
genres <- xpathSApply(xml_data, "//book/genre", xmlValue)
Books_data_XML <- data.frame(
  Title = titles,
  Authors = authors,
  Publication_Year = years,
  Summary = summaries,
  Genre = genres,
  Main_Themes = main_themes)

print(Books_data_XML)

```


##JSON

```{r}
Json_file = "https://raw.githubusercontent.com/sokkarbishoy/DATA607/main/books.JSON"
json_raw <- fromJSON(Json_file)
Books_table_json <- as.data.frame(json_raw, col.names = c(""))
Books_table_json
```


##Conclusion
In this assignment, I learned how to pull data from HTML, XML, and JSON code formate into R. I am sure this was simple as I was the one who created the table but over the next weeks I will be practicing web scrapping more and would apply what I learned from this assignment. 
