---
title: "Project 1 "
author: "Bishoy Sokkar"
date: "2024-02-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Working with Ratings text file

Text file was uploaded to personal Github account

```{r}
#load necessary packages
library(tidyverse)
library(stringr)
library(readr)

#Read text file, uploaded on personal Github

ELOratings1 = read.csv("https://raw.githubusercontent.com/sokkarbishoy/DATA607/main/tournamentinfo.txt")

ELOratings2 <- read_lines(url("https://raw.githubusercontent.com/sokkarbishoy/DATA607/main/tournamentinfo.txt"))
ELOratings3 <- str_squish(ELOratings2)
head(ELOratings3, n=15)
```

**Extracting the names of the players using REGEX**

```{r}
elorating4 <- str_view(ELOratings3, "\\s\\|\\s[A-Z].*\\s\\|\\d")
head(elorating4, n=40)
elorating4 <- str_extract(elorating4,"\\|\\s[^0-9]+\\s\\|" )
head(elorating4)
Player <- str_extract(elorating4, "\\s.*\\s")
Player <- str_trim(Player)
print(Player)
```

**Extracting States using REGEX**

```{r}



State <- str_view(ELOratings3, "^[A-Z].\\s")
head(State)
State <- str_extract(State, "[A-Z][A-Z]\\s")
print(State)



```

**Extracting total points**

```{r}
Total_points <- str_view(ELOratings3, "\\s\\|\\s[A-Z].*\\s\\|\\d.\\d")
Total_points <- str_extract(Total_points, "\\d.\\d")
print(Total_points)



```

**Extracting Pre Rating**

```{r}
Pre_rating <- str_view(ELOratings3, "R:\\s.*\\-")
Pre_rating <- str_extract(Pre_rating, ":\\s.*\\-")
Pre_rating <- str_extract(Pre_rating, "\\d.*\\d")
Pre_rating <- str_extract(Pre_rating, "\\d{3,4}")
print(Pre_rating) 
```

**Extracting players opponent average pre-rating: Still having difficulties understanding the best way to use the joint function.**

```{r}player's'}

```

build a data frame, with the extracted data, except the opponents pre rating.

```{r}
RatingsFinal <- data.frame( Player,
                            State,
                            Total_points,
                            Pre_rating)
RatingsFinal
```

Converting into a CSV

```{r}
write.csv(RatingsFinal, file = "ELO_Ratings.csv", row.names = FALSE)
```
