---
title: "Week 2 Assignment"
author: "Bishoy Sokkar"
date: "2024-02-10"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Build Table

In this assignment, we will recall tables from the SQL database into R. The data is stored in the database server: cunydata607sql.mysql.database.azure.com. The two tables created were Friends and Ratings. Friends have names, ages, occupancy, and favorite genres. Ratings contain the ratings recorded for each movie.

```{r cars}
# install packages needed
library(DBI)
library(odbc)
library(RSQLite)
library(RMySQL)
library(dplyr)

```

## 2. Store data in SQL database

The tables were made in EXCEL and uploaded through the "Table Data Import Wizard" Available on MySQL Workbench.

## 3. Transfer data from SQL database to R dataframe

In the code below, we will establish a connection with MySQL server with the stored data, and view insights from the tables.

```{r}
con <- dbConnect(MySQL(), user='bishoy.sokkar60', password='<bishoy.sokkar607>', dbname='bishoy.sokkar60', host='cunydata607sql.mysql.database.azure.com')
```

To view the tables in the database:

```{r}
dbListTables(con)
```

To learn more about the table variables in the data base:

```{r}
dbListFields(con, "friends") 
dbListFields(con, "ratings")
```

To retrieve tables from MySQL server, we use the following function:

```{r}
#to create tables in R with the data
Friends_data = dbGetQuery(con, "SELECT * FROM friends ")
Ratings_data = dbGetQuery(con, "SELECT * FROM ratings")
# to glimpse to confirm all information transfered correctly.
glimpse(Friends_data)
glimpse(Ratings_data)
```

## 4. Missing data strategy

When uploading the tables into the SQL database, I selected NN (Not Null) which constrains the column from containing null values.
